---
title: "Jackknife function in R"
author: "Team: Sweat Pants feat. Problem"
date: "March 7, 2017"
output: html_document
---

# Project Goals

The goal of this project is to work through building functions around the Jackkinfe. We will start with some simple functions and then determine a pattern to build a more general function in the end. 


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

**PHP 2560 Only** (Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)



Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r, eval=TRUE}
mean.jackknife <- function(a_vector) {
  mean.lessi <- vector()
  for(i in 1:length(a_vector)){
    mean.lessi[i] <- mean(a_vector[-i])
  }
  jackknife.stderr <- sqrt(var(mean.lessi)*(((length(mean.lessi)-1)^2)/length(mean.lessi)))  
  # Put your function in here
  return(jackknife.stderr)
}

x<- c(1:15)
mean.jackknife(x)

```



#### Jackknife for the mean Example: test your code here


```{r, eval = FALSE}
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
all.equal(formula_se_of_mean, mean.jackknife(some_normals))
```





# Project Goals

The goal of this project is to work through building functions around the Jackkinfe. We will start with some simple functions and then determine a pattern to build a more general function in the end. 


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

**PHP 2560 Only** (Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)



Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r, eval=FALSE}
mean.jackknife <- function(a_vector) {
  # Put your function in here
  return(jackknife.stderr)
}
```



#### Jackknife for the mean Example: test your code here


```{r, eval = FALSE}
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
all.equal(formula_se_of_mean,mean.jackknife(some_normals))
```



##Jackknife for Gamma Parameters


The following function is a way to calculate the method of moments estimators for the gamma distribution:

```{r, eval=TRUE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}

gamma.est(x)
```


###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval=TRUE }
gamma.jackknife <- function(a_vector) {
  mean.lessi <- vector()
  vlessi <- vector()
  alessi <- vector()
  slessi <- vector()
  
  for(i in 1:length(a_vector)){
    mean.lessi[i] <- mean(a_vector[-i])
    vlessi[i] <- var(a_vector[-i])
    alessi[i] <- mean.lessi[i]^2/vlessi[i]
    slessi[i] <- vlessi[i]/mean.lessi[i]
  }
  
  jackknife.stderrs <- c((sqrt(var(alessi)*(((length(alessi)-1)^2)/length(alessi)))),
                         (sqrt(var(slessi)*(((length(slessi)-1)^2)/length(slessi)))))
  
  return(jackknife.stderrs)
}

```



###Jackknife for Gamma Parameters Example


```{r, eval=TRUE}
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval=TRUE}
jackknife.lm <- function(df, formula, p) {
  df.lessi <- matrix(, nrow = dim(df)[1], ncol = p)
  for(i in 1:nrow(df)){
    df.lessi[i,] <- lm(formula, df[-i,])$coefficients
   }
  
  jackknife.stderr <- c((sqrt(var(df.lessi[,1])*
                                (((length(df.lessi[,1])-1)^2)/length(df.lessi[,1])))),
                         (sqrt(var(df.lessi[,2])*
                                 (((length(df.lessi[, 2])-1)^2)/length(df.lessi[, 2])))))

  return(jackknife.stderr)
}

```



###Jackknife for linear regression coefficients Example


```{r, eval=TRUE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
#Our method
jackknife.lm(df=data,formula="output~input",p=2)
```




##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 

```{r, eval=T}
 omit.case <- function(the_data, omitted_point) {
   if(is.vector(the_data)) {
     the_data <- the_data[-omitted_point]
   } else {
     the_data <- the_data[-omitted_point,]
   }
   return(the_data)
# This should take the data and omit one point at a time and return the new data
   }

```

- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  

We would need three arguments:

```{r,eval=T}
 omit_and_est <- function(omit) {
   estimator(omit.case(the_data, omitted_point = omit))
       # This function should take the output of omit.case and use it as input for the estimator
 }
```


```{r, eval=TRUE}
jackknife <- function(estimator, the_data) {
  
  # Define n to be lenth or number of rows here
  if(is.vector(the_data)) {
     n <- length(the_data)
   } else {
     n <- dim(the_data)[1]
   }
  
  omit_and_est <- function(omit) {
   estimator(omit.case(the_data, omitted_point = omit))
       # This function should take the output of omit.case and use it as input for the estimator
  }
  
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  
  var.of.reestimates <- apply(jackknife.ests, 1, var)
  
  jackknife.var <- ((n-1)^2/n)*var.of.reestimates
  
  jackknife.stderr <- sqrt(jackknife.var)
  
  return(jackknife.stderr)
}
```



---

###It works


```{r, eval=TRUE}
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)

jackknife(estimator=mean, the_data = some_normals)

all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))

```

---



```{r, eval=TRUE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input))
```

---


```{r, eval=TRUE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```


```{r, eval=TRUE}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r, eval=TRUE}
jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

<<<<<<< HEAD
```{r, eval=TRUE}
data_size <- function(the_data) {
  n <- ifelse(is.vector(the_data), length(the_data), nrow(the_data))
}
```
- Write a function that calculate the variance of all the estimates and returns the standard error
```{r, eval=TRUE}
scale_and_sqrt_vars <- function(jackknife.ests, n) {
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  var.of.reestimates <- apply(jackknife.ests, 1, var)
  jackknife.var <- ((n-1)^2/n)*var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
=======
    ```{r, eval=FALSE}
      data_size <- function(the_data) {
  
      }
    ```
- Write a function that calculate the variance of all the estimates and returns the standard error
```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jackknife.ests,n) {
  
>>>>>>> 337a20c37b2f9ce45fdbc8f0036f6d2edcf38bdf
  return(jackknife.stderr)
}
```



###Now invoke those functions

```{r, eval=TRUE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

